---
title: "ADS_Assignment"
author: "Kaixin Yu"
student ID: 1118795
output:
  html_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# please install the comment package if the packages are not installed
#install.packages("corrplot")
#install.packages("factoextra")
library('plot.matrix')
library(corrplot)
library(reshape2)
library(ggplot2)
library(matrixStats)
library("factoextra")
```

 
Question 1

Q1.1
```{r}
# create TC sources
par(mfrow=c(2,3))
par(mar=c(5.1, 2.1, 2.1, 2.1))
TC_1<- c(rep(rep(c(1,0), each =15), 8))
TC_2<- c(rep(0,20), rep(rep(c(1,0),c(20,25)), length.out=220))
TC_3 <- c(rep(rep(c(1,0), c(25, 60-25)), length.out=240))
TC_4 <- c(rep(rep(c(1,0), c(15, 40-15)), length.out=240))
TC_5 <- c(rep(rep(c(1,0), c(20, 40-20)), length.out=240))
TC_6 <- c(rep(rep(c(1,0), c(25, 40-25)), length.out=240))
TC <- matrix(c(TC_1, TC_2, TC_3, TC_4, TC_5, TC_6), 240,6)

# evaluate mean and variance of TC
colMeans(TC)
colVars(TC)

# scale TC
TC <- scale(TC)
TC_title<- c("TC_1", "TC_2", "TC_3", "TC_4", "TC_5", "TC_6")

# plot TC sources
for (i in (1:6)){
  plot(TC[,i], type='l', xlab="", ylab="", main=TC_title[i])
}

```


Q 1.2
```{r}
# plot correlation matrix of TC 
par(mfrow = c(1,1))
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(cor(TC), main="Correlation of time sources")
```

Q 1.3
```{r}
# create 6 tmpSMs and plot them out
par(mfrow=c(2,3))
tmpSM_1<- matrix(rep(0, 441), 21,21)
tmpSM_1[c(2:6), c(2:6)] <- 1
plot(tmpSM_1, border=NA,xlab="", ylab="")

tmpSM_2<- matrix(rep(0, 441), 21,21)
tmpSM_2[c(2:6), c(15:19)] <- 1
plot(tmpSM_2, border=NA, xlab="", ylab="")

tmpSM_3<- matrix(rep(0, 441), 21,21)
tmpSM_3[c(8:13), c(2:6)] <- 1
plot(tmpSM_3, border=NA, xlab="", ylab="")

tmpSM_4<- matrix(rep(0, 441), 21,21)
tmpSM_4[c(8:13), c(15:19)] <- 1
plot(tmpSM_4, border=NA, xlab="", ylab="")

tmpSM_5<- matrix(rep(0, 441), 21,21)
tmpSM_5[c(15:19), c(2:6)] <- 1
plot(tmpSM_5, border=NA, xlab="", ylab="")

tmpSM_6<- matrix(rep(0, 441), 21,21)
tmpSM_6[c(15:19), c(15:19)] <- 1
plot(tmpSM_6, border=NA, xlab="", ylab="")
```

```{r}
# plot the correlation matrix of SM
par(mfrow = c(1,1))

tmpSM <- matrix(c(as.vector(tmpSM_1), as.vector(tmpSM_2), as.vector(tmpSM_3), as.vector(tmpSM_4), as.vector(tmpSM_5), as.vector(tmpSM_6)), 6, 441, byrow = TRUE)

corrplot(cor(t(tmpSM)), addCoef.col="black", method="color")
```
Plots shows tmpSM are independent to each other.

Explore why not standardize tmpSM
```{r}
# means and variance of SM source
rowMeans(tmpSM)
rowVars(tmpSM)

# plots after standardize of SM
par(mfrow=c(2,3))
SM_standard_title <-c("SM_standard_1","M_standard_2","M_standard_3","M_standard_4","M_standard_5","M_standard_6")
for (i in 1:6){
  plot(matrix(scale(tmpSM[i,]), 21, 21), border=NA, xlab="", ylab="", main=SM_standard_title[i])
}
```
Q 1.4
```{r}
# create noise and plot correlation matrix for TC 
set.seed(30034)
par(mfrow=c(1,1))
TC_noise <- matrix(rnorm(240*6, mean=0, sd=sqrt(0.25)), 240,6)
TC_noise_cor <- cor(TC_noise)
corrplot(TC_noise_cor, method='color', addCoef.col="black", title="CM for TC noise", mar=c(0,0,1,0))

# create noise and plot correlation matrix for SM
tmpSM_noise <- matrix(rnorm(6*441, mean=0, sd=sqrt(0.015)), 6, 441)
tmpSM_noise_cor <- cor(t(tmpSM_noise))
corrplot(tmpSM_noise_cor, method='color', addCoef.col="black", title="CM for SM noise",mar=c(0,0,1,0))
```

As the plots shown, the noise of both two sources are independent.

```{r}
# plot the histogram for both noise source and add distribution curve
hist(as.vector(TC_noise),xlab = "TC noise", main="Distribution of TC noise", prob=TRUE)
curve(dnorm(x, mean=0, sd = sqrt(1.96*0.25)),col="red", lwd=2, add=TRUE )

hist(as.vector(t(tmpSM_noise)),xlab = "SM noise", main="Distribution of SM noise", prob=TRUE)
curve(dnorm(x, mean=0, sd = sqrt(1.96*0.015)), col="red", lwd=2, add=TRUE)
```

Both TC and tmpSM noise look have a normal distrbution, and satisfied the 1.96$\sigma$.

```{r}
# visualize the correlation between the noise product and observed variables
par(mfrow = c(1,1))
par(mar=c(5.1, 4.1, 4.1, 4.1))
noise_product <- TC_noise%*%tmpSM_noise
corrplot(cor(noise_product), method="color", tl.pos = 'n')
```

The correlation heap map show non correlation, noise are independent.

Q1.5
```{r}
# build X dataset
X <- (TC+TC_noise)%*%(tmpSM+tmpSM_noise)

# check the dimension of each matrix
(dim(TC))
(dim(tmpSM_noise))
(dim(TC_noise))
(dim(tmpSM))
```

The dimension of each matrix shows that TC*tmpSM_noise and TC_noise*tmpSM exists.

```{r}
# get 100 sample variables
set.seed(123)
par(mfrow = c(1,1))
par(mar=c(5.1, 4.1, 4.1, 4.1))
sample<-data.frame(n=1:240, X[, sample(ncol(X), size=100)])

# plot the sample
sample_gg <- melt(sample, id.vars="n")
ggplot(sample_gg, aes(x=n, y=value, col=variable))+ geom_line()+ labs(x="time point")
```

```{r}
# calculate variance of all variables
plot(colVars(X), xlab="variable", ylab="variance", main="Variance of all variables")

# scale X data
X<- scale(X)
```

Question 2

Q2.1
```{r}
D <- TC
# retrieve A by LSR
A_lsr <- abs(solve(t(D)%*%D)%*%t(D)%*%X)
# retrieve D by LSR
D_lsr <- X%*%t(A_lsr)
```

```{r}
# plot all retrieval sources
par(mfrow=c(1,2))
par(mar=c(5.1, 2.1, 4.1, 4.1))
A_lsr_title <-c("A_lsr_1","A_lsr_2","A_lsr_3","A_lsr_4","A_lsr_5","A_lsr_6")
D_lsr_title <- c("D_lsr_1","D_lsr_2","D_lsr_3","D_lsr_4","D_lsr_5","D_lsr_6")
for (i in 1:6){
  plot(matrix(A_lsr[i,], 21,21),border=NA,xlab="", ylab="",  main=A_lsr_title[i])
  plot(D_lsr[,i], type='l', xlab="", ylab="", main=D_lsr_title[i])
}
```

```{r}
# scatter plot between 3rd column of D_lsr and 30th column of X
par(mfrow = c(1,1))
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(D_lsr[,3], X[,30], xlab="3rd column of D_lsr", ylab="30th column of X")

# check which spatial signal stay at 1
(tmpSM[,30])
```

since only the third regressor is actived, which means the TC_3 makes contributions to 30th columns of X

Q2.2
```{r}
# select $\lambda\$ by using check and guess method
D <- TC

#randomly generate lamda
set.seed(30034)
(lambda <- runif(1,0,1))
A_rr_random <- abs(solve(t(D)%*%D + lambda*441*diag(6))%*%t(D)%*%X)
# try using other lambda =0.5
A_rr <- abs(solve(t(D)%*%D + 0.5*441*diag(6))%*%t(D)%*%X)
# estimate D with new A_rr
D_rr_random <- X%*%t(A_rr_random)
D_rr <- X%*%t(A_rr)
```

The $\lambda$ we use is 0.8990044。


```{r}
# calculate and store maximum correlation of each source
Ctlsr <- rep(NA, 6)
Ctrr_ran <- rep(NA,6)
Ctrr <- rep(NA, 6)
for (i in (1:6)){
  Ctlsr[i] <- abs(cor(D_lsr, TC))[i, i]
  Ctrr[i] <- abs(cor(D_rr, TC))[i, i]
  Ctrr_ran[i] <- abs(cor(D_rr_random, TC))[i,i]
}

Ctlsr
Ctrr
Ctrr_ran
```


```{r}
# calculate the sum of the maximum correlation score
(sum(Ctlsr))
(sum(Ctrr))
(sum(Ctrr_ran))
paste("Ctlsr is less than Ctrr_random:", sum(Ctlsr)<sum(Ctrr_ran))
paste("Ctlsr is less than Ctrr:", sum(Ctlsr)<sum(Ctrr))
paste("Ctrr_random is less than Ctrr:", sum(Ctrr_ran)<sum(Ctrr))
```

model with $\lambda$ = 0.5 performs better, use A_rr.

```{r}
# plot the absolute first vector from Arr with applied lambda= 1000
par(mfrow = c(1,1))
par(mar=c(5.1, 4.1, 4.1, 4.1))
A_rr_1000 <- abs(solve(t(D)%*%D + 1000*441*diag(6))%*%t(D)%*%X)
plot(A_lsr[1,], type='l', ylab="", main = "First A_lsr")
plot(A_rr_1000[1,], type='l',  ylab="", main = "First A_rr with lambda 1000")

```
All value in first vector of Arr shrinking towards zero.

Q2.3
```{r}
rm(.Random.seed, envir=globalenv())

# code for perform LR
MSE_lr <- array(NA, c(10, 21))
# The candidates of rho
rho <- seq(0,1,0.05)

# do 10 realization, each time with different X
for (realization in 1:10){
  if (realization==1) X_lr <-X
  else {
    # generate new noise to retireve new X
    TC_noise_lr <- matrix(rnorm(240*6, mean=0, sd=sqrt(0.25)), 240,6)
    SM_noise_lr <- matrix(rnorm(6*441, mean=0, sd=sqrt(0.015)), 6,441)
    X_lr <- (TC + TC_noise_lr)%*%(tmpSM + SM_noise_lr)
    X_lr <- scale(X_lr)
  }
  # perform Lasso algorithm for each rho
  for(rr in 1:21){
  
    step <- 1/(norm(TC %*% t(TC)) * 1.1)
    thr <- rho[rr]*240*step
    nsrcs <- 6
    Ao <- matrix(0, nsrcs, 1)
    A <- matrix(0, nsrcs, 1)
    Alr <- matrix(0, nsrcs, 21*21)

    for (k in 1:(21*21)) {
      A <- Ao+step*(t(TC) %*% (X_lr[,k]-(TC%*%Ao)))
      A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))
  
      for (i in 1:10) {
        Ao <- A
        A <- Ao+step * (t(TC)%*%(X_lr[,k]-(TC%*%Ao)))
        A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))
      }
      Alr[,k] <- A
    }
  # calculate MSE of each rho
    Alr <- abs(Alr)
    Dlr <- X_lr%*%t(Alr)
    MSE <- sum((X_lr-Dlr%*%Alr)**2)/(240*441)
    MSE_lr[realization, rr] <- MSE
  }
}
```

```{r}
# calculate the mean MSE and plot out
avg_MSE_lr <- colMeans(MSE_lr)
plot(rho, avg_MSE_lr, main="the avergae MSE of each rho")
```

```{r}
(avg_MSE_lr)
# get the location of rho that has minimum MSE
index <- which.min(avg_MSE_lr)
paste("The minimum average MSE is",  avg_MSE_lr[index])

# get rho value
rho_lr <- rho[index]
paste("The value of rho with minimum MSE is",rho_lr)
```

At the minimum MSE, \rho = 0.6, the value of average MSE is 0.7659844

Q2.4
```{r}
# use rho_lr = 0.6, estimate A_lr parameter
step <- 1/(norm(TC %*% t(TC)) * 1.1)
thr <- rho_lr*240*step
Ao <- matrix(0, nsrcs, 1)
A <- matrix(0, nsrcs, 1)
A_lr <- matrix(0, nsrcs, 21*21)

for (k in 1:(21*21)) {
  A <- Ao+step*(t(TC) %*% (X[,k]-(TC%*%Ao)))
  A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))
  
  for (i in 1:10) {
    Ao <- A
    A <- Ao+step * (t(TC)%*%(X[,k]-(TC%*%Ao)))
    A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))
  }
  A_lr[,k] <- A
}

# use A_lr to estimate D_lr
A_lr <- abs(A_lr)
D_lr <- X%*%t(A_lr)
```

```{r}
# compute all four correlation vector
Ctrr <- rep(NA, 6)
Csrr <- rep(NA, 6)
Ctlr <- rep(NA, 6)
Cslr <- rep(NA, 6)
for (i in (1:6)){
  Ctrr[i] <- abs(cor(D_rr, TC))[i, i]
  Csrr[i] <- abs(cor(t(A_rr), t(tmpSM)))[i, i]
  Ctlr[i] <- abs(cor(D_lr, TC))[i, i]
  Cslr[i] <- abs(cor(t(A_lr), t(tmpSM)))[i, i]

}

Ctrr
Csrr
Ctlr
Cslr
```

```{r}
# sum of four correlation vector
(sum_Ctrr<-sum(Ctrr))
(sum_Csrr<-sum(Csrr))
(sum_Ctlr<-sum(Ctlr))
(sum_Cslr<-sum(Cslr))

paste("Ctrr is less than Ctlr:", sum_Ctrr<sum_Ctlr)
paste("Csrr is less than Cslr:", sum_Csrr<sum_Cslr)
```

sum of Ctlr and Cslr are both larger than Ctrr and Csrr

```{r}
# plot D_rr, A_rr, D_lr, A_lr
par(mfrow=c(6,4))
par(mar=c(1,1,1,1))
plot(matrix(A_rr[1,], 21,21), border=NA,xlab="", ylab="",key=NULL,
     main="A_Ridge")

plot(D_rr[,1], type='l', xlab="", ylab="", main="D_Ridge")

plot(matrix(A_lr[1,], 21,21), border=NA,xlab="", ylab="",key=NULL,
     main="A_Lasso")
plot(D_lr[,1], type='l', xlab="", ylab="", main="D_Lasso")

for (i in 2:6){
  plot(matrix(A_rr[i,], 21,21), border=NA,xlab="", ylab="", main="",key=NULL)
  plot(D_rr[,i], type='l', xlab="", ylab="",main="")

  plot(matrix(A_lr[i,], 21,21), border=NA,xlab="", ylab="",main="",key=NULL)
  plot(D_lr[,i], type='l', xlab="", ylab="",main="")
}

```


Q2.5
```{r}
# build PCR model
PC <- prcomp(D, scale.=TRUE)

# get eigenvalue and plot them out
(eig_pc <- PC$sdev^2)
fviz_eig(PC, geom="line", choice = "eigenvalue", addlabels = TRUE)+ labs(title = "Eigenvalue-PCA", x = "Principal Components", y = "Eigenvalue")

```

The 6th PC owns the minimum eigenvalue, 0.2. 

```{r}
par(mfrow=c(1,2))
par(mar=c(5.1, 2.1, 4.1, 4.1))

# retrive Z from PCR model
Z_pc <- PC$x

# plot all Z
Z_title<- c("Z_PC_1", "Z_PC_2", "Z_PC_3", "Z_PC_4", "Z_PC_5", "Z_PC_6")

for (i in 1:6){
  plot(Z_pc[,i], type='l', xlab="", ylab="", main=Z_title[i])
  plot(TC[,i], type='l', xlab="", ylab="",main=TC_title[i])
}
```
```{r}
plot(cor(Z_pc), main="Correlation of time sources")
```

```{r}
# perform lasso with Z on X
step <- 1/(norm(Z_pc %*% t(Z_pc)) * 1.1)
rho_pcr <- 0.001
thr <- rho_pcr*240*step
Ao <- matrix(0, nsrcs, 1)
A <- matrix(0, nsrcs, 1)
A_pcr <- matrix(0, nsrcs, 21*21)

for (k in 1:(21*21)) {
  A <- Ao+step*(t(Z_pc) %*% (X[,k]-(Z_pc%*%Ao)))
  A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))
  
  for (i in 1:10) {
    Ao <- A
    A <- Ao+step * (t(Z_pc)%*%(X[,k]-(Z_pc%*%Ao)))
    A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))
  }
  A_pcr[,k] <- A
}

# use A_lr to estimate D_lr
A_pcr <- abs(A_pcr)
D_pcr <- X%*%t(A_pcr)
```

```{r}
# plot A_pcr and D_pcr
par(mfrow=c(1,2))
par(mar=c(5.1, 3.1, 4.1, 4.1))
Apcr_title<- c("A_PCR_1", "A_PCR_2", "A_PCR_3", "A_PCR_4", "A_PCR_5", "A_PCR_6")
Dpcr_title<- c("D_PCR_1", "D_PCR_2", "D_PCR_3", "D_PCR_4", "D_PCR_5", "D_PCR_6")
for (i in 1:6){
  plot(matrix(A_pcr[i,], 21,21), border=NA,xlab="", ylab="", 
        main=Apcr_title[i])
  plot(D_pcr[,i], type='l', xlab="", ylab="",main=Dpcr_title[i])
}
```
```{r}
# evaluate the MSE of all four model
(MSE_lsr <- sum((X-D_lsr%*%A_lsr)**2)/(240*441))
(MSE_rr <- sum((X-D_rr%*%A_rr)**2)/(240*441))
(MSE_lr <- sum((X-D_lr%*%A_lr)**2)/(240*441))
(MSE_pcr <- sum((X-D_pcr%*%A_pcr)**2)/(240*441))

MSE_all <- data.frame(MSE_lsr= sum((X-D_lsr%*%A_lsr)**2)/(240*441),
                      MSE_rr=sum((X-D_rr%*%A_rr)**2)/(240*441), 
                      MSE_lr = sum((X-D_lr%*%A_lr)**2)/(240*441),
                      MSE_pcr=sum((X-D_pcr%*%A_pcr)**2)/(240*441))
bar<- barplot(as.matrix(MSE_all))
#text(x=bar, y=as.matrix(MSE_all), labels = as.matrix(MSE_all),  pos = 3, cex = 0.8)
```

```{r}
# evalaute which regressor contributes most for the first two PC
para <- get_pca_var(PC)
plot(Z_pc[,1], Z_pc[,2])
fviz_contrib(PC, choice = "var", axes = 1:2)
```
The 5th, 6th and 4th contributes most in first two PCs.

```{r}
# compare spatial maps across four models
par(mfrow=c(6,4))
par(mar=c(1,1,1,1))

#plot all four model retrived A
plot(matrix(abs(A_lsr[1,]), 21,21), border=NA,xlab="", ylab="",key=NULL,
     main="A_LSR")

plot(matrix(abs(A_rr[1,]), 21,21), border=NA,xlab="", ylab="",key=NULL,
     main="A_Ridge") 

plot(matrix(abs(A_lr[1,]), 21,21), border=NA,xlab="", ylab="",key=NULL,
     main="A_Lasso")

plot(matrix(abs(A_pcr[1,]), 21,21), border=NA,xlab="", ylab="",key=NULL,
     main="A_PCA")

for (i in 2:6){
  plot(matrix(abs(A_lsr[i,]), 21,21), border=NA,xlab="", ylab="",key=NULL,
     main="")

plot(matrix(abs(A_rr[i,]), 21,21), border=NA,xlab="", ylab="",key=NULL,
     main="")

plot(matrix(abs(A_lr[i,]), 21,21), border=NA,xlab="", ylab="",key=NULL,
     main="")

plot(matrix(abs(A_pcr[i,]), 21,21), border=NA,xlab="", ylab="",key=NULL,
     main="")
}
```

```{r}
# compare temporal souces across four models
par(mfrow=c(6,4))
par(mar=c(1,1,1,1))

#plot all four model retrived A
plot(D_lsr[,1], type='l', xlab="", ylab="", main="D_LSR")
plot(D_rr[,1], type='l', xlab="", ylab="", main="D_Ridge")
plot(D_lr[,1], type='l', xlab="", ylab="", main="D_Lasso")
plot(D_pcr[,1], type='l', xlab="", ylab="", main="D_PCR")

for (i in 2:6){
  plot(D_lsr[,i], type='l', xlab="", ylab="", main="")
  plot(D_rr[,i], type='l', xlab="", ylab="", main="")
  plot(D_lr[,i], type='l', xlab="", ylab="", main="")
  plot(D_pcr[,i], type='l', xlab="", ylab="", main="")

}
```
